**REPRODUCIBLE RESEARCH**

**ASSIGNMENT 1**

**Mar 15, 2015**

# Loading and preprocessing the data

## Load the data
```{r}
library(ggplot2)
Sys.setlocale('LC_TIME', 'English') # Because I am using a Chinese OS.
data <- read.csv('activity.csv', header = TRUE)
```

## Process/transform the data (if necessary) into a format suitable for your analysis

All NA data are transformed into zeroes. And all values of total steps walked at any particular time-point are calculated.
```{r}
data$total_steps[1] = 0
for (i in 1:nrow(data)) {
    if (is.na(data$steps[i])) {
        data$steps_no[i] = 0
    } else {
        data$steps_no[i] = data$steps[i]
    }
}
for (i in 2:nrow(data)) {
    data$total_steps[i] = data$total_steps[i-1] + data$steps_no[i]
}
```

# What is mean total number of steps taken per day?

Dates are transformed into numbers of day.
```{r}
for (i in 1:nrow(data)) {
    data$day[i] = as.numeric(data$date[i])
}
```

## Total numbers of steps per day are calculated and stored into a new dataset.
```{r}
total_days = max(data$day)
daily_steps = data.frame(1:total_days)
colnames(daily_steps) = "day"
for (i in 1:total_days) {
    temp_data = subset(data, day == i)
    daily_steps$sum[i] <- sum(temp_data$steps)
}
```

## Total numbers of steps per day are calculated.
```{r}
print(daily_steps)
```

## A histogram of the total number of steps taken each day is created.
```{r}
hist(daily_steps$sum)
ggplot(daily_steps, aes(x = sum)) + geom_histogram()
```

## Calculating the mean and median total number of steps taken per day
```{r}
total_steps <- as.matrix(summary(daily_steps$sum))
mean = total_steps[4,1]
median = total_steps[3,1]
```
Mean of the total number of steps taken per day is `r mean`.
Median of the total number of steps taken per day is `r median`.

# What is the average daily activity pattern?

## A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

Average number of steps taken across all days within 5-minute intervals is calcuated.
```{r}
max_interval = max(data$interval)
interval = seq(0, max_interval, 5)
daily_interval = data.frame(interval)
for (i in 0:(max_interval/5)) {
    temp_interval = subset(data, interval == i*5)
    temp_interval$steps[is.na(temp_interval$steps)] <- 0
    daily_interval$sum[i+1] <- sum(temp_interval$steps)
}
daily_interval$avg = daily_interval$sum / total_days
```

A time series plot is made.
```{r}
plot(x = daily_interval$interval, y = daily_interval$avg, type = 'l')
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
sorted <- daily_interval[order(daily_interval$avg),]
max_day <- sorted[nrow(sorted), 1]
```
No. `r max_day` interval contains the maximum number of steps.


# Imputing missing values

## Calculate and report the total number of missing values in the dataset
```{r}
data.NA <- read.csv('activity.csv', header = TRUE)
total_no_NA = nrow(subset(data.NA, is.na(steps)))
```
Total number of missing values in the dataset is `r total_no_NA`.

## Devise a strategy for filling in all of the missing values in the dataset
```{r}
data$steps = data$steps_no
```
Note: Missing values have already been filled early on.

## Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
new.data <- data[, 1:3]
```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
daily_steps_no_NA = data.frame(1:total_days)
colnames(daily_steps_no_NA) = "day"
for (i in 1:total_days) {
    temp_data = subset(data, day == i)
    daily_steps_no_NA$sum[i] <- sum(temp_data$steps)
}
print(daily_steps_no_NA)
hist(daily_steps_no_NA$sum)
summary(daily_steps_no_NA$sum)
total_steps_no_NA <- as.matrix(summary(daily_steps_no_NA$sum))
mean_no_NA = total_steps_no_NA[4,1]
median_no_NA = total_steps_no_NA[3,1]
```
**After filling missing values:**
Mean of the total number of steps taken per day is `r mean_no_NA`.
Median of the total number of steps taken per day is `r median_no_NA`.
Both are different from those obtained from original data that contains NA data.



# Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
new.data$wd <- weekdays(as.Date(new.data$date, abbreviate = TRUE))
for (i in 1:nrow(new.data)) {
    if (new.data$wd[i] == 'Saturday' | new.data$wd[i] == 'Sunday') {
        new.data$wd_class[i] = 'WE'
    }
    else {
        new.data$wd_class[i] = 'WD'
    }
}
```

## Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
new.data.WE <- subset(new.data, wd_class == 'WE')
new.data.WD <- subset(new.data, wd_class == 'WD')

new.data.WE_interval = data.frame(interval)
for (i in 0:(max_interval/5)) {
    temp_data_WE = subset(new.data.WE, interval == i*5)
    new.data.WE_interval$sum[i+1] <- sum(temp_data_WE$steps)
}
new.data.WE_interval$avg = new.data.WE_interval$sum / total_days / 2

new.data.WD_interval = data.frame(interval)
for (i in 0:(max_interval/5)) {
    temp_data_WD = subset(new.data.WD, interval == i*5)
    new.data.WD_interval$sum[i+1] <- sum(temp_data_WD$steps)
}
new.data.WD_interval$avg = new.data.WD_interval$sum / total_days / 5

par(mar = c(2, 1, 1, 1))
par(mfrow = c(2, 1))

WE <- plot(x = new.data.WE_interval$interval, y = new.data.WE_interval$avg, type = 'l')
WD <- plot(x = new.data.WD_interval$interval, y = new.data.WD_interval$avg, type = 'l')
```

There are differences in activity patterns between weekdays and weekends.
